### 2016-02-19 hans/bayes-tardis-luminosity-prediction.tex
[hans] Add frequency depends, update assumptions

\begin{mtemize}
\item Spectrum of luminosities in channels (or ``bins'')
  $b=0,1,\ldots,B$ is given by the telescope
\item In the simplest case, the frequency of channel $b$ is given by
  $\nu_b = \nu_0 + b\,\delta\nu$, where $\delta\nu$ is
  the bin width. In the general case, the bins are not of equal
  size. Let's then consider $\nu_b$ as the left edge of bin $b$,
  and $\nu_{B+1}$ is the right edge of the last bin; i.e., the
  maximum frequency.
\item We treat the results of a single TARDIS run as \textit{data}
  $D$, i.e.\ fixed results which we do not question but accept as
  given.  These ``data'' are the number of packets leaving the
  supernova, $\npack$, the luminosity $\bml$, and the frequencies
  $\bmnu$ for each packet
  \begin{align}
    \lleq{dfc}
    D = \{ \npack, \bml, \bmnu \} \,,
  \end{align}
  where
  \begin{align}
    \lleq{dfd}
    % \bml_b &= (\ell_{b,1},\ell_{b,2},\ldots,\ell_{b,n_b})
    \dim \bml = \dim \bmnu = \npack \,.
  \end{align}
  Given $D$, the number of samples in bin $b$ is
  \begin{align}
    \lleq{dff}
    n_b = \sum_{j=1}^{\npack} \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j )
  \end{align}
  and similarly the luminosity is
  \begin{align}
    \lleq{dfe}
    \ell_b = \sum_{j=1}^{\npack}\ell_j \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j ),
  \end{align}
  where $\mathbf{1}_V(x)$ is the indicator function requiring $x \in V$.
\end{mtemize}

@@

Our goal of prediction of $L$ given the data $(n,\bml)$ is captured in
the probability $p(L\cond n,\bml)$ with all relevant parameters
integrated out. We first note from Eq.~(\ref{pmh}) that $p(N\cond
n,\bml) = p(N\cond n)$ does not depend on $\bml$. Furthermore, since
$L$ is fully determined by $(N,\bmL)$,
\begin{align}
  \lleq{pmjd}
  L &= \sum_{j=1}^N L_j
\end{align}
is does not depend directly on $(n,\bml)$, i.e.\
\begin{align}
  \lleq{pmi}
  p(L\cond N,\bmL,n,\bml)
  = p(L\cond N,\bmL)
  &= \delta(L - \textstyle\sum_{j=1}^N L_j).
\end{align}
\checked{}
Hence the prediction is given by
\begin{align}
  \lleq{pmj}
  p(L\cond n,\bml)
  &= \sum_{N=0}^\infty p(L\cond N,n,\bml)\,p(N\cond n,\bml) \\
  \lleq{pmjb}
  &= \sum_N p(N\cond n)\,\int d\bmL\, p(L\cond N,\bmL,n,\bml)\, p(\bmL\cond N,n,\bml)
  \\
  \lleq{pmjc}
  &= \sum_N p(N\cond n)\,\int d\bmL\,\delta(L - \textstyle\sum_{j=1}^N L_j)
  \, p(\bmL\cond N,n,\bml)
\end{align}
\checked{}
The dependence on the TARDIS ``data'' $(n,\bml)$ enters through
$p(\bmL\cond N,n,\bml)$ in terms of the parameters $\bmphi$ of
$p(\bmL\cond N,\bmphi)$,
\begin{align}
  \lleq{pmk}
  p(\bmL\cond N,n,\bml)
  &= \int d\bmphi\, p(\bmL\cond N,\bmphi,n,\bml)\,p(\bmphi\cond N,n,\bml) \\
  &= \int d\bmphi\, p(\bmL\cond N,\bmphi)\,p(\bmphi\cond N,n,\bml).
\end{align}
because the likelihood for $\bmL$ is fully determined by $N,\bmphi$
and hence independent of $(n,\bml)$.
%
Again using Bayes' Theorem (\ref{bth})
\begin{align}
  \lleq{pmkb}
  p(\bmphi\cond N,n,\bml) &=
  \frac{p(\bml\cond N,n,\bmphi)\,p(\bmphi\cond N,n)} %
  {\int d\bmphi\,p(\bml\cond N,n,\bmphi)\,p(\bmphi\cond N,n)} %
  % \ =\ \frac{p(\bml\cond N,n,\bmphi)\,p(\bmphi\cond N,n)} %
  % {p(\bml\cond n)} %
\end{align}
\fred{$p(\bmphi\cond N,n,\bml) = p(\bmphi\cond n,\bml)$. Below $N$ is (correctly) missing in \refeq{pmp} }
and using a multiplicity-independent prior for $\bmphi$
\begin{align}
  \lleq{pml}
  p(\bmphi\cond N,n) &= p(\bmphi)
\end{align}
we can write (\ref{pmk}) as
\begin{align}
  \lleq{pmn}
  p(\bmL\cond N,n,\bml)
  &= \int d\bmphi\, p(\bmL\cond N,\bmphi)\, %
  \frac{p(\bml\cond n,\bmphi)\,p(\bmphi)} %
  {\int d\bmphi\,p(\bml\cond n,\bmphi)\,p(\bmphi)} \\
  \lleq{pmo}
  &= \frac{1}  {p(\bml\cond n)} %
  \int d\bmphi\, p(\bmL\cond N,\bmphi)\, %
  p(\bml\cond n,\bmphi)\,p(\bmphi)
\end{align}
with $p(\bml\cond n) = \int d\bmphi\, p(\bml\cond
n,\bmphi)\,p(\bmphi)$ in the denominator.
Inserting this into (\ref{pmjc}), we obtain
\begin{equation}
  \lleq{pmp}
  \boxed{
  p(L\cond n,\bml)
  = \frac{1}{p(\bml\cond n)}
  \sum_N p(N\cond n)\,\int d\bmL\,\delta(L - \textstyle\sum_j L_j)
  \displaystyle\int d\bmphi\,p(\bmL\cond N,\bmphi) \, p(\bml\cond n,\bmphi)
  \, p(\bmphi)
  }
\end{equation}
\checked{}
with
\begin{align}
  p(\bmL\cond N,\bmphi) &= \prod_{j=1}^N p(L_j\cond \bmphi) \\
  p(\bml\cond n,\bmphi) &= \prod_{j=1}^n p(\ell_j\cond \bmphi)
\end{align}
\fred{And $p(\cdot \cond \bmphi)$ is the same function in both equations.}
In other words: Once we have decided on a prior $p(\bmphi)$ and a
specific pdf to model the likelihood $p(L_j\cond \bmphi)$, we have a
prediction for $L$ which takes into account all possible $N$,
$\lambda$ and $\bmphi$.

>>>>

\begin{mtemize}
\item Spectrum of luminosities in channels (or ``bins'')
  $b=0,1,\ldots,B$ is given by the telescope
\item In the simplest case, the frequency of channel $b$ is given by
  $\nu_b = \nu_0 + b\,\delta\nu$, where $\delta\nu$ is
  the bin width. In the general case, the bins are not of equal
  size. Let's then consider $\nu_b$ as the left edge of bin $b$,
  and $\nu_{B+1}$ is the right edge of the last bin; i.e., the
  maximum frequency.
  \item We treat the results of a single TARDIS run as \textit{data}
  \fred{If in the end we want to compare to actual data from a
    telescope, I'd prefer to not call the TARDIS output ``data''. Why
    don't we just call it ``TARDIS output''?}  $D$, i.e.\ fixed
  results which we do not question but accept as given.  These
  ``data'' are the number of packets leaving the supernova, $\npack$,
  the luminosity $\bml$, and the frequencies $\bmnu$ for each packet
  \begin{align}
    \lleq{dfc}
    D = \{ \npack, \bml, \bmnu \} \,,
  \end{align}
  where
  \begin{align}
    \lleq{dfd}
    % \bml_b &= (\ell_{b,1},\ell_{b,2},\ldots,\ell_{b,n_b})
    \dim \bml = \dim \bmnu = \npack \,.
  \end{align}
  Given $D$, the number of samples in bin $b$ is
  \begin{align}
    \lleq{dff}
    n_b = \sum_{j=1}^{\npack} \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j )
  \end{align}
  and similarly the luminosity is
  \begin{align}
    \lleq{dfe}
    \ell_b = \sum_{j=1}^{\npack}\ell_j \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j ),
  \end{align}
  where $\mathbf{1}_V(x)$ is the indicator function requiring $x \in V$.
\end{mtemize}

@@

Our goal of predicting $L$ given the TARDIS output $(\npack,\bml,
\bmnu)$ is captured in the probability $p(L\cond \npack,\bml, \bmnu)$
with all relevant parameters integrated out. We first note from
\refeq{pmh} that $p(N\cond \npack,\bml, \bmnu) = p(N\cond n)$ does not
depend on $\bml$ or $\bmnu$ directly but only through $n$, the number
of packets in the bin; cf. \refeq{dff}. Furthermore, since $L$ is
fully determined by $(N,\bmL)$,
\begin{align}
  \lleq{pmjd}
  L &= \sum_{j=1}^N L_j
\end{align}
is does not depend directly on $(\npack),\bml,\bmnu)$, i.e.\
\begin{align}
  \lleq{pmi}
  p(L\cond N,\bmL,n,\bml, \bmnu)
  = p(L\cond N,\bmL)
  &= \delta(L - \textstyle\sum_{j=1}^N L_j).
\end{align}
Hence the prediction is given by
\begin{align}
  \lleq{pmj}
  p(L\cond n,\bml, \bmnu)
  &= \sum_{N=0}^\infty p(L\cond N,n,\bml, \bmnu)\,p(N\cond n,\bml, \bmnu) \\
  \lleq{pmjb}
  &= \sum_N p(N\cond n)\,\int d\bmL\, p(L\cond N,\bmL,n,\bml, \bmnu)\, p(\bmL\cond N,n,\bml, \bmnu)
  \\
  \lleq{pmjc}
  &= \sum_N p(N\cond n)\,\int d\bmL\,\delta(L - \textstyle\sum_{j=1}^N L_j)
  \, p(\bmL\cond N,n,\bml, \bmnu)
\end{align}
\fred{Add missing $\bmnu$! Everything that is conditional on $\bml$ is
  also conditional on $\bmnu$} The dependence on the TARDIS ``data''
$(\npack,\bml, \bmnu)$ enters through $p(\bmL\cond N,\npack,\bml,
\bmnu)$ in terms of the parameters $\bmphi$ of $p(\bmL\cond
N,\bmphi)$,
\begin{align}
  \lleq{pmk}
  p(\bmL\cond N,\npack,\bml, \bmnu)
  &= \int d\bmphi\, p(\bmL\cond N,\bmphi,\npack,\bml, \bmnu)\,p(\bmphi\cond N,\npack,\bml, \bmnu) \\
  &= \int d\bmphi\, p(\bmL\cond N,\bmphi)\,p(\bmphi\cond N,\npack,\bml, \bmnu).
\end{align}
because the likelihood for $\bmL$ is fully determined by $N,\bmphi$
and hence independent of $(\npack,\bml, \bmnu)$.
%
Again using Bayes' Theorem (\ref{bth})
\begin{align}
  \lleq{pmkb}
  p(\bmphi\cond N,\npack,\bml, \bmnu) &=
  \frac{p(\bml\cond N,\npack,\bmphi,\bmnu)\,p(\bmphi\cond N,\npack,\bmnu)} %
  {\int d\bmphi\,p(\bml\cond N,\npack,\bmphi,\bmnu)\,p(\bmphi\cond N,\npack,\bmnu)}, %
  % \ =\ \frac{p(\bml\cond N,\npack,\bmphi)\,p(\bmphi\cond N,n)} %
  % {p(\bml\cond n)} %
\end{align}

 a multiplicity-independent prior for $\bmphi$
\begin{align}
  \lleq{pml}
  p(\bmphi\cond N,\npack,\bmnu) &= p(\bmphi),
\end{align}
and recognizing that $N$ does not tell us anything about $\bml$
\begin{align}
  \lleq{pmla}
  p(\bml\cond N,\npack,\bmphi,\bmnu) = p(\bml\cond \npack,\bmphi,\bmnu)
\end{align}
we can write (\ref{pmk}) as
\begin{align}
  \lleq{pmn}
  p(\bmL\cond N,\npack,\bml, \bmnu)
  &= \int d\bmphi\, p(\bmL\cond N,\bmphi)\, %
  \frac{p(\bml\cond \npack,\bmphi)\,p(\bmphi)} %
  {\int d\bmphi\,p(\bml\cond \npack,\bmphi)\,p(\bmphi)} \\
  \lleq{pmo}
  &= \frac{1}  {p(\bml\cond n)} %
  \int d\bmphi\, p(\bmL\cond N,\bmphi)\, %
  p(\bml\cond \npack,\bmphi)\,p(\bmphi)
\end{align}
with $p(\bml\cond n) = \int d\bmphi\, p(\bml\cond
\npack,\bmphi)\,p(\bmphi)$ in the denominator.
Inserting this into (\ref{pmjc}), we obtain
\begin{equation}
  \lleq{pmp}
  \boxed{
  p(L\cond \npack,\bml, \bmnu)
  = \frac{1}{p(\bml\cond n)}
  \sum_N p(N\cond n)\,\int d\bmL\,\delta(L - \textstyle\sum_j L_j)
  \displaystyle\int d\bmphi\,p(\bmL\cond N,\bmphi) \, p(\bml\cond \npack,\bmphi)
  \, p(\bmphi)
  }
\end{equation}
with
\begin{align}
  p(\bmL\cond N,\bmphi) &= \prod_{j=1}^N p(L_j\cond \bmphi) \\
  p(\bml\cond \npack,\bmphi) &= \prod_{j=1}^n p(\ell_j\cond \bmphi)
\end{align}
\fred{And $p(\cdot \cond \bmphi)$ is the same function in both equations.}
In other words: Once we have decided on a prior $p(\bmphi)$ and a
specific pdf to model the likelihood $p(L_j\cond \bmphi)$, we have a
prediction for $L$ which takes into account all possible $N$,
$\lambda$ and $\bmphi$.
